Analysis paper

### 2021-04-20 analysis/analysis.Rmd
update plots in report

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(patchwork)

# load and set up data
cases <- combined_data[target_type == "case" & 
                         as.Date(forecast_date) %in% forecast_dates$cases]
deaths <- combined_data[target_type == "death" & 
                         as.Date(forecast_date) %in% forecast_dates$deaths]
data <- rbindlist(list(cases, deaths))
```

@@

@@

truth_vs_forecast_2 <- df %>%
  ggplot(aes(y = true_value, x = target_end_date, colour = model, fill = model)) + 
  geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
                 size = 1.5,
                 alpha = 0.4,
                 position = position_dodge(width = 3)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 0.8,
                 position = position_dodge(width = 3)) + 
  labs(y = "value", x = "date") + 
  geom_point() + 
  geom_line() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom")

@@

```{r fig.height=10, fig.width=15}
# plot with scores over time ---------------------------------------------------
scores_target_forecastdate <- eval_forecasts(
  data = data[grepl("2", target)],
  summarise_by = c("model", "location_name", "target_type", "forecast_date"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

wis_scores_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date")

@@

```{r}
### same plot with mean score ratio
rel_skill_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = scaled_rel_skill, fill = model, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "relative skill", x = "date")

ggsave(here("analysis", "plots", "wis-over-time.png"), 
       plot = rel_skill_time_2)
```

```{r fig.height=16, fig.width=8}
# glue two plots together
truth_vs_forecast_2 / rel_skill_time_2 + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom') 
ggsave(here("analysis", "plots", "figure-forecasts-2.png"))
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)

# load and set up data
cases <- combined_data[target_type == "case" & 
                         as.Date(forecast_date) %in% forecast_dates$cases]
deaths <- combined_data[target_type == "death" & 
                         as.Date(forecast_date) %in% forecast_dates$deaths]
data <- rbindlist(list(cases, deaths))
data[, location_target := paste0(target_type, "s", " in ", location_name)]
```

## Visualisation of data

truth <- dcast(df, ... ~ quantile, value.var = "prediction")[
  , .(location_name, target_end_date, target_type, true_value)] %>%
  unique()

truth <- truth[order(target_type, location_name, target_end_date)]

# classify epidemic according to whether, on a given forecast date, the two
# weeks before that have seen monotonic rise, decline, or an unclear trend
classify_epidemic <- function(data, cutoff = 0.05, growth_cutoff = 2) {
  dt <- as.data.table(data)

  # calculate differences and set differences smaller than 
  # a certain cutoff to zero
  dt[, diff_1 := c(NA, diff(true_value, 1)), 
     by = c("location_name", "target_type")]
  dt[abs(diff_1) < (cutoff * true_value), diff_1 := 0]
  dt[, diff_prev := c(shift(diff_1, 1)), 
     by = c("location_name", "target_type")]

  # assign a label depending on observed differences
  dt[, c("classification", "speed") := "unclear"]
  dt[(diff_1 >= 0 & diff_prev >= 0), 
     classification := "increasing"]
  dt[(diff_1 <= 0 & diff_prev <= 0 ), 
     classification := "decreasing"]
  dt[(diff_1 == 0 & diff_prev == 0), 
     classification := "unclear"]
  dt[(classification == "increasing") & 
       diff_1 > growth_cutoff * diff_prev, 
     speed := "accelerating"]
    dt[(classification == "increasing") & 
         diff_1 < 1/growth_cutoff * diff_prev, 
     speed := "decelerating"]
  dt[(classification == "decreasing") & 
       abs(diff_1) > growth_cutoff * abs(diff_prev), 
     speed := "accelerating"]
  dt[(classification == "decreasing") & 
       abs(diff_1) < 1/growth_cutoff * abs(diff_prev), 
     speed := "decelerating"]

  dt[, c("diff_1", "diff_prev") := NULL]
  return(dt)
}

dt <- classify_epidemic(truth)

p <- ggplot(dt, aes(x = target_end_date)) +

  geom_line(aes(y = true_value)) + 
  geom_point(aes(y = true_value, color = classification)) + 
  labs(y = "value", x = "date") + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom")

show_acceleration <- FALSE
if (show_acceleration) {
  p <- p + 
    geom_rect(aes(xmin = target_end_date - 7,
                  xmax = target_end_date,
                  ymin = -Inf,
                  ymax = Inf,
                  fill = speed), alpha = 0.2) +
    scale_fill_manual(values = c(
      "accelerating" = "darkgrey",
      "decelerating" = "lightgrey",
      "unclear" = "white"
    ))
}

print(p)

@@

```{r fig.height=10, fig.width=15}
df <- combined_data[as.Date(forecast_date) %in% forecast_dates$unfiltered &
                      grepl("2", target) & 
                      quantile %in% c(0.025, 0.25, 0.75, 0.975)]
df[, location_target := paste0(target_type, "s", " in ", location_name)]

@@

df <- classify_epidemic(df)

truth_vs_forecast_2 <- df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  # geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
  #                size = 1.5,
  #                alpha = 0.4,
  #                position = position_dodge(width = 3)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 1,
                 position = position_dodge(width = 6)) + 
  labs(y = "value", x = "date") + 
  geom_point(aes(shape = classification)) + 
  geom_line() + 
  theme_light() + 
  facet_wrap(~ location_target, scales = "free", ncol = 1) + 
  theme(legend.position = "bottom") + 
  scale_color_brewer(palette = "Accent") + 
  theme(panel.background = element_rect(fill = "#e1f0fe"))

@@

```{r fig.height=10, fig.width=15}
# plot with scores over time ---------------------------------------------------
scores_target_forecastdate <- eval_forecasts(
  data = data[grepl("2", target)],
  summarise_by = c("model", "location_name", "target_type", "forecast_date", "location_target"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)
truth <- classify_epidemic(truth)
tmp <- copy(truth)
tmp[, target_end_date := as.Date(target_end_date) + 2]
scores_target_forecastdate[, forecast_date := as.Date(forecast_date)]
setnames(tmp, old = "target_end_date", new = "forecast_date")
scores_target_forecastdate <- merge(
  scores_target_forecastdate, tmp, 
  by = c("location_name", "forecast_date", "target_type"), 
  all.x = TRUE)

wis_scores_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_wrap(~ location_target, scales = "free", ncol = 1) +
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date") +
  scale_color_brewer(palette = "Accent") + 
  theme(panel.background = element_rect(fill = "#e1f0fe"))

@@

```{r}
### same plot with mean score ratio
rel_skill_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = scaled_rel_skill, fill = model, colour = model)) + 
  geom_line() + 
  geom_point(aes(shape = classification)) + 
  theme_light() + 
  facet_wrap( ~ location_target, ncol = 1, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "relative skill", x = "date") +
  scale_color_brewer(palette = "Accent") + 
  theme(panel.background = element_rect(fill = "#e1f0fe"))

ggsave(here("analysis", "plots", "rel-skill-over-time.png"), 
       plot = rel_skill_time_2)
```

```{r fig.height=10, fig.width=15}
# glue two plots together
truth_vs_forecast_2 | rel_skill_time_2 + 
  plot_layout(guides = "collect") & 
  theme(legend.position = 'bottom') 
ggsave(here("analysis", "plots", "figure-forecasts-2.png"))
```

@@

palette = 
**To do**
